---
title: "Emmeans alternative - PP66 FM3.0"
author: "Valerie Martin"
date: "10/7/2019"
output:
  word_document: default
  html_document: default
---

## Skypilot Selection Analyse with 66% Pollination Percentage and 3.0 Seed Maturity


```{r include=FALSE}
setwd("~/Desktop/Skypilot 2017")
library(ggplot2)
library(plyr)
library(nlme)
library(car)
library(emmeans)
library(magrittr)
library(multcompView)

fruit <- read.csv("Skypilot 2017 (MASTER) Tab3 08-12-19.csv", header=T) # all data included (final)
fruit$OtherSeedcount[is.na(fruit$OtherSeedcount)] <- 0
fruit$Total.SeedsPerPlant <- fruit$No.TrmtSeeds+fruit$OtherSeedcount

cols <- c("Site", "Habitat", "Treatment", "Plant.Number", "Morph", "No.Pollinated.Total", "Total.RepUnits", "Prct.Pollinated",
          "Avg.SepL",	"Avg.TubeL",	"Avg.LobeL",	"Avg.LobW",	"Avg.CorollaFlare", "No.SeedsPerFruit", "Fruit.Maturity.Rank", 
          "Total.SeedsPerFruit", "Total.SeedsPerPlant")
fruit.sm <- subset(fruit, select=cols) # remove raw data columns

fruit.sm <- subset(fruit.sm, Avg.CorollaFlare>0) # remove those without flower measurements
fruit.sm$Prct.Pollinated <- as.numeric(as.character(fruit.sm$Prct.Pollinated)) # change Prct.Pollinated to numeric (it was a factor)
fruit.sm <- subset(fruit.sm, !(Treatment=="Supplementation" & Prct.Pollinated < 0.666)) # remove those supplement flowers with <75% fruits pollinated
fruit.sm <- subset(fruit.sm, Fruit.Maturity.Rank < 3) # remove fruit maturity 2.5 (includes all exlusion as well) 
fruit.sm$CorollaL <- fruit.sm$Avg.TubeL + fruit.sm$Avg.LobeL
max(fruit.sm$Total.SeedsPerPlant)
fruit.sm$RelativeFitness <- fruit.sm$Total.SeedsPerPlant/138


fruit.sw <- subset(fruit.sm, Morph=="Sweet") # subst to sweet morph only
fruit.sw.np <- subset(fruit.sw, Site %in% c("Niwot Ridge", "Penn Mountain"))
fruit.sw.nat <- subset(fruit.sw, Treatment=="Natural")
fruit.sw.sup <- subset(fruit.sw, Treatment=="Supplementation")
fruit.nat <- subset(fruit.sm, Treatment=="Natural") # subset to natural treatment only
fruit.sup <- subset(fruit.sm, Treatment=="Supplementation") # subset to supplement treatment only
fruit.np <- subset(fruit.sm, Site %in% c("Niwot Ridge", "Penn Mountain"))

flwr <- subset(fruit, Avg.CorollaFlare>0) # remove those without flower measurements
flwr.nat <- subset(flwr, Treatment=="Natural") # subset to natural treatment only
```

### Supplementation Experiment
Niki: "Can't use cumberland pass (no supplmement seeds were successfully collected...triple checking that).  Marginal interaction effect: total seeds per plant increased with supplementation of skunky, but not sweet, flowers. Sweet has this weird relationship where sup seedset goes up with flower size. Jake swears size differences shouldn't have influenced the effectiveness of the treatment."

#### Sweet morph only 
Niki: "no effect of habitat => remove
no treatment effect, but Candi saw this in some years - might it relate to weaking selection?"

```{r eval=FALSE, include=FALSE}
qqnorm(fruit.sw.np$Total.SeedsPerPlant)
qqline(fruit.sw.np$Total.SeedsPerPlant)
hist(fruit.sw.np$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.sw.np$Total.SeedsPerPlant))
qqline(sqrt(fruit.sw.np$Total.SeedsPerPlant))
hist(sqrt(fruit.sw.np$Total.SeedsPerPlant))
```

Niki: "Extract least squares mean which accounts for random effect"
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment, random=~1|Site, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit)

#type="response" causes the emmeans and contrasts tables to backtransform from the sqrt and show confidence intervals and p-values that represent the backtransformed variables
marginal = emmeans(fit, specs=pairwise ~ Treatment, type = "response")
marginal$emmeans
marginal$contrasts

marginal$contrasts %>%
     summary(infer = TRUE)

#putting the data into a data frame so that we can plot it
marginal_contrasts = marginal$contrasts %>%
     summary() %>%
     as.data.frame()
marginal_contrasts

marginal_sqrt = emmeans(fit, specs=pairwise ~ Treatment)
#plot of the sqrt transformed emmeans
plot(marginal_sqrt, xlab="Estimated marginal means")
```

Plots (not controlling for random effects)
```{r echo=TRUE}
ggplot(fruit.sw.np, aes(Treatment,Total.SeedsPerPlant)) +
  geom_violin(fill='gray') +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)  +
  facet_wrap(~Morph) +
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  xlab("Morph") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))

ggplot(fruit.sw.np, aes(Treatment,Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  xlab("Morph") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))
```

With habitat effects: 
Val: "For the two second ANOVA tables, I am getting an error regarding the use of lm() instead of lme() - why are we using lm() and why is it not computing into an ANOVA table?"
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment, random=~1|Site/Habitat, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lm(sqrt(Total.SeedsPerPlant)~Habitat, data=fruit.sw.np)

fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Habitat, random=~1|Site, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lm(sqrt(Total.SeedsPerPlant)~Site, data=fruit.sw.np)

```


Both Morphs
```{r include=FALSE}
qqnorm(fruit.np$Total.SeedsPerPlant)
qqline(fruit.np$Total.SeedsPerPlant)
hist(fruit.np$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.np$Total.SeedsPerPlant))
qqline(sqrt(fruit.np$Total.SeedsPerPlant))
hist(sqrt(fruit.np$Total.SeedsPerPlant))
```

```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Morph, random=~1|Site, data=fruit.np, method="ML", na.action=na.omit)  
anova.lme(fit)

marginal_both = emmeans(fit, specs=pairwise ~ Treatment, type = "response")
marginal_both$emmeans
marginal_both$contrasts
marginal_both_sqrt = emmeans(fit, specs=pairwise ~ Treatment)
marginal_both$contrasts %>%
     summary(infer = TRUE)

#putting the data into a data frame so that we can plot it
marginal_contrasts2 = marginal_both$contrasts %>%
     summary() %>%
     as.data.frame()
marginal_contrasts2 


#plot of the sqrt transformed emmeans
plot(marginal_both_sqrt, xlab="Estimated marginal means")
```

plot (not controlling for random effects)
```{r echo=TRUE}
ggplot(fruit.sm, aes(Treatment,Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  facet_wrap(~Morph) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))
```

with habitat effects: 
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Habitat, random=~1|Site, data=fruit.np, method="ML", na.action=na.omit)  
anova.lme(fit)
```
