---
title: "Selection Analyses 11.29"
author: "Valerie Martin"
date: "11/29/2019"
output:
  word_document: default
  html_document: default
---

## Skypilot Selection Analyses with 75% Pollination Percentage and 2.5 Seed Maturity
### Looking at combined PCA for natural flowers and Total Seeds/Plant by Size*Morph
#### (directly comparing these analyses to those from the stigma analyses)

#### Loading data and creating subsets...


```{r include=FALSE}
setwd("~/Desktop/Skypilot 2017/Skypilot Selection")
library(ggplot2)
library(plyr)
library(nlme)
library(car)
library(emmeans)
library(magrittr)
library(multcompView)

fruit <- read.csv("Skypilot 2017 (MASTER) Tab3 08-12-19.csv", header=T) # all data included (final)
fruit$OtherSeedcount[is.na(fruit$OtherSeedcount)] <- 0
fruit$Total.SeedsPerPlant <- fruit$No.TrmtSeeds+fruit$OtherSeedcount

cols <- c("Site", "Habitat", "Treatment", "Plant.Number", "Morph", "No.Pollinated.Total", "Total.RepUnits", "Prct.Pollinated",
          "Avg.SepL",	"Avg.TubeL",	"Avg.LobeL",	"Avg.LobW",	"Avg.CorollaFlare", "No.SeedsPerFruit", "Fruit.Maturity.Rank", 
          "Total.SeedsPerFruit", "Total.SeedsPerPlant")
fruit.sm <- subset(fruit, select=cols) # remove raw data columns

fruit.sm <- subset(fruit.sm, Avg.CorollaFlare>0) # remove those without flower measurements
fruit.sm$Prct.Pollinated <- as.numeric(as.character(fruit.sm$Prct.Pollinated)) # change Prct.Pollinated to numeric (it was a factor)
fruit.sm <- subset(fruit.sm, !(Treatment=="Supplementation" & Prct.Pollinated < 0.75)) # remove those supplement flowers with <75% fruits pollinated
fruit.sm <- subset(fruit.sm, Fruit.Maturity.Rank < 2.5) # remove fruit maturity 2.5 (includes all exlusion as well) 
fruit.sm$CorollaL <- fruit.sm$Avg.TubeL + fruit.sm$Avg.LobeL
max(fruit.sm$Total.SeedsPerPlant)
fruit.sm$RelativeFitness <- fruit.sm$Total.SeedsPerPlant/138


fruit.sw <- subset(fruit.sm, Morph=="Sweet") # subst to sweet morph only
fruit.sw.np <- subset(fruit.sw, Site %in% c("Niwot Ridge", "Penn Mountain"))
fruit.sw.nat <- subset(fruit.sw, Treatment=="Natural")
fruit.sw.sup <- subset(fruit.sw, Treatment=="Supplementation")
fruit.nat <- subset(fruit.sm, Treatment=="Natural") # subset to natural treatment only
fruit.sup <- subset(fruit.sm, Treatment=="Supplementation") # subset to supplement treatment only
fruit.np <- subset(fruit.sm, Site %in% c("Niwot Ridge", "Penn Mountain"))

flwr <- subset(fruit, Avg.CorollaFlare>0) # remove those without flower measurements
flwr.nat <- subset(flwr, Treatment=="Natural") # subset to natural treatment only
```


### Total Seeds/Plant by Size Measurement*Morph
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit, type="marginal")
summary(fit)

fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.TubeL*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
```


#### Flower measurements combined via pca (all 4 traits - same as pressed) - BOTH MORPHS
```{r}
measures <- c("Avg.CorollaFlare", "Avg.TubeL", "Avg.LobeL", "Avg.LobW")
flwr.size<-as.matrix(subset(fruit.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.nat))
##### Almost all loadings for PC2 are negative, but Avg. TubeL has a very high positive
fruit.nat$PC1.all <- pc.extracted$PC1 
fruit.nat$PC2.all <- pc.extracted$PC2

```


#### Total Seeds/Plant by PC1*Morph and PC2*Morph
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~PC1.all*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)

fit<-lme(sqrt(Total.SeedsPerPlant)~PC2.all*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
```