---
title: "Skypilot Selection Analyses__PP_75 - SM_2.5 Part 2"
author: "Valerie Martin"
date: "8/25/2019"
output:
  word_document: 
    fig_height: 2.25
    fig_width: 3
  html_document:
    df_print: paged
---

## Skypilot Selection Analyse with 75% Pollination Percentage and 2.5 Seed Maturity

#### Loading data and creating subsets...


```{r include=FALSE}
setwd("~/Desktop/Skypilot 2017/Skypilot Selection")
library(ggplot2)
library(plyr)
library(nlme)
library(car)
library(emmeans)
library(magrittr)
library(multcompView)

fruit <- read.csv("Skypilot 2017 (MASTER) Tab3 08-12-19.csv", header=T) # all data included (final)
fruit$OtherSeedcount[is.na(fruit$OtherSeedcount)] <- 0
fruit$Total.SeedsPerPlant <- fruit$No.TrmtSeeds+fruit$OtherSeedcount

cols <- c("Site", "Habitat", "Treatment", "Plant.Number", "Morph", "No.Pollinated.Total", "Total.RepUnits", "Prct.Pollinated",
          "Avg.SepL",	"Avg.TubeL",	"Avg.LobeL",	"Avg.LobW",	"Avg.CorollaFlare", "No.SeedsPerFruit", "Fruit.Maturity.Rank", 
          "Total.SeedsPerFruit", "Total.SeedsPerPlant")
fruit.sm <- subset(fruit, select=cols) # remove raw data columns

fruit.sm <- subset(fruit.sm, Avg.CorollaFlare>0) # remove those without flower measurements
fruit.sm$Prct.Pollinated <- as.numeric(as.character(fruit.sm$Prct.Pollinated)) # change Prct.Pollinated to numeric (it was a factor)
fruit.sm <- subset(fruit.sm, !(Treatment=="Supplementation" & Prct.Pollinated < 0.75)) # remove those supplement flowers with <75% fruits pollinated
fruit.sm <- subset(fruit.sm, Fruit.Maturity.Rank < 2.5) # remove fruit maturity 2.5 (includes all exlusion as well) 
fruit.sm$CorollaL <- fruit.sm$Avg.TubeL + fruit.sm$Avg.LobeL
max(fruit.sm$Total.SeedsPerPlant)
fruit.sm$RelativeFitness <- fruit.sm$Total.SeedsPerPlant/138


fruit.sw <- subset(fruit.sm, Morph=="Sweet") # subst to sweet morph only
fruit.sw.np <- subset(fruit.sw, Site %in% c("Niwot Ridge", "Penn Mountain"))
fruit.sw.nat <- subset(fruit.sw, Treatment=="Natural")
fruit.sw.sup <- subset(fruit.sw, Treatment=="Supplementation")
fruit.nat <- subset(fruit.sm, Treatment=="Natural") # subset to natural treatment only
fruit.sup <- subset(fruit.sm, Treatment=="Supplementation") # subset to supplement treatment only
fruit.np <- subset(fruit.sm, Site %in% c("Niwot Ridge", "Penn Mountain"))

flwr <- subset(fruit, Avg.CorollaFlare>0) # remove those without flower measurements
flwr.nat <- subset(flwr, Treatment=="Natural") # subset to natural treatment only
```

### Supplementation Experiment
Niki: "Can't use cumberland pass (no supplmement seeds were successfully collected...triple checking that).  Marginal interaction effect: total seeds per plant increased with supplementation of skunky, but not sweet, flowers. Sweet has this weird relationship where sup seedset goes up with flower size. Jake swears size differences shouldn't have influenced the effectiveness of the treatment."

#### Sweet morph only 
Niki: "no effect of habitat => remove
no treatment effect, but Candi saw this in some years - might it relate to weaking selection?"

```{r eval=FALSE, include=FALSE}
qqnorm(fruit.sw.np$Total.SeedsPerPlant)
qqline(fruit.sw.np$Total.SeedsPerPlant)
hist(fruit.sw.np$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.sw.np$Total.SeedsPerPlant))
qqline(sqrt(fruit.sw.np$Total.SeedsPerPlant))
hist(sqrt(fruit.sw.np$Total.SeedsPerPlant))
```

Niki: "Extract least squares mean which accounts for random effect"
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment, random=~1|Site, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit)

#type="response" causes the emmeans and contrasts tables to backtransform from the sqrt and show confidence intervals and p-values that represent the backtransformed variables
marginal = emmeans(fit, specs=pairwise ~ Treatment, type = "response")
marginal$emmeans
marginal$contrasts

marginal$contrasts %>%
     summary(infer = TRUE)

#putting the data into a data frame so that we can plot it
marginal_contrasts = marginal$contrasts %>%
     summary() %>%
     as.data.frame()
marginal_contrasts

#plot of the sqrt transformed emmeans
plot(marginal_contrasts, xlab="Estimated marginal means")
```

Plots (not controlling for random effects)
```{r echo=TRUE}
ggplot(fruit.sw.np, aes(Treatment,Total.SeedsPerPlant)) +
  geom_violin(fill='gray') +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)  +
  facet_wrap(~Morph) +
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  xlab("Morph") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))

ggplot(fruit.sw.np, aes(Treatment,Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  xlab("Morph") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))
```

With habitat effects: 
Val: "For the two second ANOVA tables, I am getting an error regarding the use of lm() instead of lme() - why are we using lm() and why is it not computing into an ANOVA table?"
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment, random=~1|Site/Habitat, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit, type="marginal")
fit<-lm(sqrt(Total.SeedsPerPlant)~Habitat, data=fruit.sw.np)
anova(fit)
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Habitat, random=~1|Site, data=fruit.sw.np, method="ML", na.action=na.omit)  
anova.lme(fit, type="marginal")
fit<-lm(sqrt(Total.SeedsPerPlant)~Site, data=fruit.sw.np)
anova(fit)
```


Both Morphs
```{r include=FALSE}
qqnorm(fruit.np$Total.SeedsPerPlant)
qqline(fruit.np$Total.SeedsPerPlant)
hist(fruit.np$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.np$Total.SeedsPerPlant))
qqline(sqrt(fruit.np$Total.SeedsPerPlant))
hist(sqrt(fruit.np$Total.SeedsPerPlant))
```

```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Morph, random=~1|Site, data=fruit.np, method="ML", na.action=na.omit)  
anova.lme(fit)

marginal_both = emmeans(fit, specs=pairwise ~ Treatment, type = "response")
marginal_both$emmeans
marginal_both$contrasts
marginal_both_sqrt = emmeans(fit, specs=pairwise ~ Treatment)
marginal_both$contrasts %>%
     summary(infer = TRUE)

#putting the data into a data frame so that we can plot it
marginal_contrasts2 = marginal_both$contrasts %>%
     summary() %>%
     as.data.frame()
marginal_contrasts2 


#plot of the sqrt transformed emmeans
plot(marginal_both_sqrt, xlab="Estimated marginal means")
```

plot (not controlling for random effects)
```{r echo=TRUE}
ggplot(fruit.sm, aes(Treatment,Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_continuous(trans="sqrt") +
  ylab("Seeds per plant") +
  facet_wrap(~Morph) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=18))
```

with habitat effects: 
```{r echo=TRUE}
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Habitat, random=~1|Site, data=fruit.np, method="ML", na.action=na.omit)  
anova.lme(fit)
```

### Selection - sweet flowers only
##### Would experience the strongest selection and sk flowers biased by habitat (almost exclusively in the krummholz).
##### Including Cumberland pass - replication is low but not biased.

#### Seeds per plant
```{r}
qqnorm(fruit.sw.nat$Total.SeedsPerPlant)
qqline(fruit.sw.nat$Total.SeedsPerPlant)
hist(fruit.sw.nat$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.sw.nat$Total.SeedsPerPlant))
qqline(sqrt(fruit.sw.nat$Total.SeedsPerPlant))
hist(sqrt(fruit.sw.nat$Total.SeedsPerPlant))
```

#### Site
```{r}
anova(lm(sqrt(Total.SeedsPerPlant)~Site, data=fruit.sw.nat, na.action=na.omit))
TukeyHSD(aov(sqrt(Total.SeedsPerPlant)~Site, data=fruit.sw.nat, na.action=na.omit))
ggplot(fruit.sw.nat, aes(x=Site,y=Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_sqrt()+
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
```

#### Corolla Flare
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=Avg.CorollaFlare,y=Total.SeedsPerPlant, color=Site)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  ylab("Seeds per plant") +
  xlab("Corolla flare") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20), legend.position = "top")
fit<-lm(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare*Site, data=fruit.sw.nat, na.action=na.omit)  
Anova(fit, type="III")

#Corolla Flare by morph
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit, type="marginal")
```

#### Tube Length
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.TubeL, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=Avg.TubeL,y=Total.SeedsPerPlant, color=Site)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  ylab("Seeds per plant") +
  xlab("Tube length") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20), legend.position = "top")
fit<-lm(sqrt(Total.SeedsPerPlant)~Avg.TubeL*Site, data=fruit.sw.nat, na.action=na.omit)  
Anova(fit, type="III")
```

### Flower measurements combined via PCA
#### Flower measurements combined via PCA - Tube Length
```{r}
cor(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$Avg.TubeL)
cor.test(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$Avg.TubeL)
plot(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$Avg.TubeL)
measures <- c("Avg.CorollaFlare", "Avg.TubeL")
flwr.size<-as.matrix(subset(fruit.sw.nat, select=measures))
pc<-princomp(flwr.size)
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.sw.nat))
fruit.sw.nat$PC1 <- pc.extracted$PC1
```

#### Flower measurement combined via PCA - Corolla Length
```{r}
fruit.sw.nat$CorollaL <- fruit.sw.nat$Avg.TubeL + fruit.sw.nat$Avg.LobeL
cor(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$CorollaL)
cor.test(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$CorollaL)
plot(fruit.sw.nat$Avg.CorollaFlare,fruit.sw.nat$CorollaL)

measures <- c("CorollaL", "Avg.CorollaFlare")
flwr.size<-as.matrix(subset(fruit.sw.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.sw.nat))
fruit.sw.nat$PC1.cl <- pc.extracted$PC1
```

#### Flower measurements combined via pca (all 4 traits - same as pressed)
```{r}
measures <- c("Avg.CorollaFlare", "Avg.TubeL", "Avg.LobeL", "Avg.LobW")
flwr.size<-as.matrix(subset(fruit.sw.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.sw.nat))
##### Inverted PC1 when added to the df, because all loadings were negative.  This way, larger PC denotes a larger flower.
fruit.sw.nat$PC1.all <- -pc.extracted$PC1 
```

#### Total seedset per plant (annual fecundity)
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~PC1.all, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=PC1,y=Total.SeedsPerPlant, color=Site)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  ylab("Seeds per plant") +
  xlab("PC1") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20), legend.position = "top")

fit<-lme(sqrt(Total.SeedsPerPlant)~PC1.cl, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
```

#### Relative fecundity (standardized by max annual fecundity)
```{r}
max(fruit.sw.nat$Total.SeedsPerPlant)
fruit.sw.nat$RelativeFitness <- fruit.sw.nat$Total.SeedsPerPlant/138

qqnorm(fruit.sw.nat$RelativeFitness)
qqline(fruit.sw.nat$RelativeFitness)
hist(fruit.sw.nat$RelativeFitness)
qqnorm(sqrt(fruit.sw.nat$RelativeFitness))
qqline(sqrt(fruit.sw.nat$RelativeFitness))
hist(sqrt(fruit.sw.nat$RelativeFitness))

fit<-lme(sqrt(RelativeFitness)~PC1, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=PC1,y=RelativeFitness)) +
  geom_point(size=2) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("black")) +
  ylab("Relative fitness") +
  xlab("PC1") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

fit<-lme(sqrt(RelativeFitness)~PC1.cl, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=PC1.cl,y=RelativeFitness)) +
  geom_point(size=2) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("black")) +
  xlim(-1,1) +
  ylab("Relative fitness") +
  xlab("PC1") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

fit<-lme(sqrt(RelativeFitness)~Avg.CorollaFlare, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=CorollaL,y=RelativeFitness)) +
  geom_point(size=2) + 
  scale_y_sqrt()+
  ylab("Relative fitness") +
  xlab("Corolla flare (cm)") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

fit<-lme(sqrt(RelativeFitness)~CorollaL, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=CorollaL,y=RelativeFitness)) +
  geom_point(size=2) + 
  scale_y_sqrt()+
  ylab("Relative fitness") +
  xlab("Corolla length (cm)") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

fit<-lme(sqrt(RelativeFitness)~Avg.TubeL, random=~1|Site, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
ggplot(fruit.sw.nat, aes(x=Avg.TubeL,y=RelativeFitness)) +
  geom_point(size=2) + 
  scale_y_sqrt()+
  ylab("Relative fitness") +
  xlab("Flower tube length (cm)") +  
  geom_hline(yintercept=0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
```


### With supplementation treatment
##### (treatment was more effective for larger flowers...Jake swears flwr size didn't affect the trmt)

#### Total seedset per plant
```{r}
qqnorm(fruit.sw$Total.SeedsPerPlant)
qqline(fruit.sw$Total.SeedsPerPlant)
hist(fruit.sw$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.sw$Total.SeedsPerPlant))
qqline(sqrt(fruit.sw$Total.SeedsPerPlant))
hist(sqrt(fruit.sw$Total.SeedsPerPlant))

fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
ggplot(fruit.sw, aes(x=Avg.CorollaFlare,y=Total.SeedsPerPlant, color=Treatment, shape=Treatment)) +
  scale_shape_manual(values=c(16,17)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("blue", "red")) +
  geom_smooth(method='lm', fill=NA) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.sw.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.sw.sup, method="ML", na.action=na.omit)  
anova.lme(fit)

fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Avg.TubeL, random=~1|Site/Habitat, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*CorollaL, random=~1|Site/Habitat, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
```

#### Relative fitness
```{r}
fit<-lme(sqrt(RelativeFitness)~Treatment*Avg.CorollaFlare, random=~1|Site, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lme(sqrt(RelativeFitness)~Treatment*Avg.CorollaFlare, random=~1|Site, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lme(sqrt(RelativeFitness)~Treatment*Avg.CorollaFlare, random=~1|Site, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
```

#### Seeds per fruit
```{r}
qqnorm(fruit.sw$No.SeedsPerFruit)
qqline(fruit.sw$No.SeedsPerFruit)
hist(fruit.sw$No.SeedsPerFruit)
qqnorm(sqrt(fruit.sw$No.SeedsPerFruit))
qqline(sqrt(fruit.sw$No.SeedsPerFruit))
hist(sqrt(fruit.sw$No.SeedsPerFruit))

fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)

ggplot(fruit.sw, aes(x=Avg.CorollaFlare,y=No.SeedsPerFruit, color=Treatment, shape=Treatment)) +
  scale_shape_manual(values=c(16,17)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("blue", "red")) +
  geom_smooth(method='lm', fill=NA) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))


fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.TubeL, random=~1|Site/Habitat, data=fruit.sw, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)

fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.LobW, random=~1|Site/Habitat, data=fruit.sw, method="ML")  
anova.lme(fit)
summary(fit)

fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.LobeL, random=~1|Site/Habitat, data=fruit.sw, method="ML")  
anova.lme(fit)
summary(fit)
```

### Both morphs: natural treatment only
#### Seeds per plant
##### Without supplementation treatment
```{r}
qqnorm(fruit.nat$Total.SeedsPerPlant)
qqline(fruit.nat$Total.SeedsPerPlant)
hist(fruit.nat$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.nat$Total.SeedsPerPlant))
qqline(sqrt(fruit.nat$Total.SeedsPerPlant))
hist(sqrt(fruit.nat$Total.SeedsPerPlant))

####Site
anova(lm(sqrt(Total.SeedsPerPlant)~Morph, data=fruit.nat, na.action=na.omit))
anova(lm(sqrt(Total.SeedsPerPlant)~Site, data=fruit.nat, na.action=na.omit))
TukeyHSD(aov(sqrt(Total.SeedsPerPlant)~Site, data=fruit.nat, na.action=na.omit))
ggplot(fruit.nat, aes(x=Site,y=Total.SeedsPerPlant)) +
  geom_boxplot() + 
  scale_y_sqrt()+
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
#corolla flare
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare*Morph, random=~1|Site, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
#tube length
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.TubeL*Morph, random=~1|Site, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
```

#### Flower measurements combined via pca (all 4 traits - same as pressed)
```{r}
measures <- c("Avg.CorollaFlare", "Avg.TubeL", "Avg.LobeL", "Avg.LobW")
flwr.size<-as.matrix(subset(fruit.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.nat))
fruit.nat$PC1.all <- pc.extracted$PC1 
# annual fecundity
fit<-lme(sqrt(Total.SeedsPerPlant)~PC1.all*Morph, random=~1|Site, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)

ggplot(fruit.nat, aes(x=Avg.CorollaFlare,y=No.SeedsPerFruit, color=Morph, shape=Treatment)) +
  scale_shape_manual(values=c(16,17)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("blue", "red")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
```

### All morphs: site, habitat and morph are random effects...
#### Seeds per fruit
```{r}
qqnorm(fruit.sm$No.SeedsPerFruit)
qqline(fruit.sm$No.SeedsPerFruit)
hist((fruit.sm$No.SeedsPerFruit))
hist(sqrt(fruit.sm$No.SeedsPerFruit))
qqnorm(sqrt(fruit.sm$No.SeedsPerFruit))
qqline(sqrt(fruit.sm$No.SeedsPerFruit))

fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.CorollaFlare, random=~1|Site/Habitat/Morph, data=fruit.sm, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit) # sign. dif by treatment but difficult to interpret figures, if transformed looks like both go up with corolla flare and only intercept is different...
fruit.nat <- subset(fruit.sm, Treatment=="Natural")
fruit.sup <- subset(fruit.sm, Treatment=="Supplementation")

fit<-lme(sqrt(No.SeedsPerFruit)~Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.nat, method="ML")
anova.lme(fit)
fit<-lme(sqrt(No.SeedsPerFruit)~Avg.CorollaFlare, random=~1|Site/Habitat, data=fruit.sup, method="ML")
anova.lme(fit)

fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*Avg.TubeL, random=~1|Site/Habitat/Morph, data=fruit.sm, method="ML")  
anova.lme(fit)
summary(fit)

fit<-lme(sqrt(No.SeedsPerFruit)~Avg.LobeL, random=~1|Site/Habitat/Morph, data=fruit.nat, method="ML")  
anova.lme(fit)
summary(fit)


ggplot(fruit.sm, aes(x=Avg.CorollaFlare,y=No.SeedsPerFruit, color=Treatment, shape=Treatment)) +
  geom_point(size=4) + 
  scale_colour_manual(values=c("blue", "red")) +
  xlim(0.8,2.4) +
  ylim(0,15) +
  geom_smooth(method='lm', aes(linetype=Treatment), fill=NA) +
  ylab("Number of seeds per fruit") +
  xlab("Corolla Flare (cm)") +
  geom_hline(yintercept=0, linetype="dashed") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
ggplot(fruit.sm, aes(x=Avg.CorollaFlare,y=No.SeedsPerFruit, color=Treatment, shape=Treatment)) +
  geom_point(size=4) + 
  scale_y_continuous(limits=c(0,15), trans="sqrt") +
  xlim(0.8,2.4) +
  scale_colour_manual(values=c("blue", "red")) +
  geom_smooth(method='lm', aes(linetype=Treatment), fill=NA) +
  ylab("Number of seeds per fruit") +
  xlab("Corolla Flare (cm)") +
  geom_hline(yintercept=0, linetype="dashed") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

ggplot(fruit.sm, aes(x=Avg.TubeL,y=No.SeedsPerFruit, color=Treatment, shape=Treatment)) +
  geom_point(size=4) + 
  scale_y_continuous(limits=c(0,15), trans="sqrt") +
  scale_colour_manual(values=c("blue", "red")) +
  ylab("Number of seeds per fruit") +
  xlab("Tube Length (cm)") +  geom_hline(yintercept=0, linetype="dashed") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))

ggplot(fruit.sm, aes(x=Avg.LobeL,y=No.SeedsPerFruit, color=Treatment, shape=Treatment)) +
  geom_point(size=4) + 
  scale_y_sqrt()+
  scale_colour_manual(values=c("blue", "red")) +
  geom_smooth(method='lm', fill=NA) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
```

### Add tube length and lobe length to get corolla length
```{r}
fruit.sm$CorollaL <- fruit.sm$Avg.TubeL + fruit.sm$Avg.TubeL
fit<-lme(sqrt(No.SeedsPerFruit)~Treatment*CorollaL, random=~1|Site/Habitat/Morph, data=fruit.sm, method="ML")  
anova.lme(fit)
summary(fit)
```

#### Seeds per plant
```{r}
qqnorm(fruit.sm$Total.SeedsPerPlant)
qqline(fruit.sm$Total.SeedsPerPlant)
hist(fruit.sm$Total.SeedsPerPlant)
qqnorm(sqrt(fruit.sm$Total.SeedsPerPlant))
qqline(sqrt(fruit.sm$Total.SeedsPerPlant))
hist(sqrt(fruit.sm$Total.SeedsPerPlant))

fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Avg.CorollaFlare, random=~1|Site/Habitat/Morph, data=fruit.sm, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit) 

fit<-lme(sqrt(Total.SeedsPerPlant)~Treatment*Avg.TubeL, random=~1|Site/Habitat/Morph, data=fruit.sm, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
```




### FLOWER SIZE by habitat and morph
#### Notes from Niki's last analysis: 
##### no difference in flower size by scent morph or habitat => sweet and skunky have the same size flowers
###### flowers are cumberland pass are smaller than niwot ridge & penn mountain. 
##### This parallels differences in seed set, but when we test for a direct relationship via regression (see below), there is no difference
##### Perhaps it has more to do with a weather effect, since selection doesn't vary among sites??? [stilll need to verify this last statement]

```{r}
qqnorm(flwr.nat$Avg.CorollaFlare)
qqline(flwr.nat$Avg.CorollaFlare)
hist(flwr.nat$Avg.CorollaFlare)

fit<-lm(Avg.CorollaFlare~Site, data=flwr.nat, na.action=na.omit)  
anova(fit)
TukeyHSD(aov(Avg.CorollaFlare~Site, data=flwr.nat, na.action=na.omit))
ggplot(flwr.nat, aes(x=Site,y=Avg.CorollaFlare)) +
  geom_boxplot() + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
fit<-lme(Avg.CorollaFlare~Site, random=~1|Morph, data=flwr.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lme(Avg.CorollaFlare~Habitat*Morph, random=~1|Site, data=flwr.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lm(Avg.CorollaFlare~Site*Morph, data=flwr.nat, na.action=na.omit)  
Anova(fit,type="III") 
fit<-lm(Avg.CorollaFlare~Site*Morph*Habitat, data=flwr.nat, na.action=na.omit)  
anova(fit) # unbalanced, so cannot run type 3. 

qqnorm(flwr.nat$Avg.TubeL)
qqline(flwr.nat$Avg.TubeL)
hist(flwr.nat$Avg.TubeL)
fit<-lm(Avg.TubeL~Site, data=flwr.nat, na.action=na.omit)  
anova(fit)
fit<-lme(Avg.TubeL~Habitat*Morph, random=~1|Site, data=flwr.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lm(Avg.TubeL~Site*Morph, data=flwr.nat, na.action=na.omit)  
Anova(fit,type="III") 
fit<-lm(Avg.TubeL~Site*Morph*Habitat, data=flwr.nat, na.action=na.omit)  
anova(fit) # unbalanced, so cannot run type 3. 

measures <- c("Avg.CorollaFlare", "Avg.TubeL", "Avg.LobeL", "Avg.LobW")
flwr.size<-as.matrix(subset(flwr.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=flwr.nat))
flwr.nat$PC1.all <- pc.extracted$PC1 
qqnorm(flwr.nat$PC1.all)
qqline(flwr.nat$PC1.all)
fit<-lm(PC1.all~Site, data=flwr.nat, na.action=na.omit)  
anova(fit)
ggplot(flwr.nat, aes(x=Site,y=PC1.all)) +
  geom_boxplot() + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text = element_text(size=20))
fit<-lme(PC1.all~Habitat*Morph, random=~1|Site, data=flwr.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
fit<-lm(PC1.all~Site*Morph, data=flwr.nat, na.action=na.omit)  
Anova(fit,type="III") 
```


### Total Seeds/Plant by Size Measurement*Morph
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.CorollaFlare*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit, type="marginal")
summary(fit)

fit<-lme(sqrt(Total.SeedsPerPlant)~Avg.TubeL*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
```


#### Flower measurements combined via pca (all 4 traits - same as pressed) - BOTH MORPHS
```{r}
measures <- c("Avg.CorollaFlare", "Avg.TubeL", "Avg.LobeL", "Avg.LobW")
flwr.size<-as.matrix(subset(fruit.nat, select=measures))
pc<-prcomp(flwr.size,center=TRUE)
print(pc)
plot(pc, type="l")
summary(pc)
pc.extracted <- as.data.frame(predict(pc, newdata=fruit.nat))
##### Almost all loadings for PC2 are negative, but Avg. TubeL has a very high positive
fruit.nat$PC1.all <- pc.extracted$PC1 
fruit.nat$PC2.all <- pc.extracted$PC2

```


#### Total Seeds/Plant by PC1*Morph and PC2*Morph
```{r}
fit<-lme(sqrt(Total.SeedsPerPlant)~PC1.all*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)

fit<-lme(sqrt(Total.SeedsPerPlant)~PC2.all*Morph, random=~1|Site/Habitat, data=fruit.nat, method="ML", na.action=na.omit)  
anova.lme(fit)
summary(fit)
```



```{r}
library(psych)
data <- fruit.nat$No.SeedsPerFruit
describe(data)

data <- fruit.sw.nat$No.SeedsPerFruit
describe(data)

data <- fruit.sk.nat

```